---
title: "Homework 2"
author: "Xinyi Wang"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
trashwheel_df = read_xlsx(
  "~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)

```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 1

