---
title: "Homework 2"
author: "Xinyi Wang"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
trashwheel_df = read_xlsx(
  "~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
)

```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"~/R_columbia/p8105/p8105_hw2_xw2758/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

First, I read and clean the NYC Transit data.

```{r read&clean, message = F, warning=F}
subway_data = read_csv("~/R_columbia/p8105/p8105_hw2_xw2758/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

**Briefly explain**

+ This dataset is about NYC subway information, including station name, station latitude / longitude, routes served, entry, vending, entrance type, and so on.

+ I just use janitor to make colnames tidy and select some interesting columns. Also, I turn some character variable into logical variable by using recode function.

+ This dataset has `r nrow(subway_data)` rows and `r ncol(subway_data)` columns. And it is not tidy.

**Answer some questions by using these data**

```{r}
distinct_station = distinct(subway_data, line, station_name, .keep_all = TRUE)
```

+ There are `r nrow(distinct_station)` distinct stations.

+  `r nrow(filter(subway_data, ada == TRUE))` stations are ADA compliant.

```{r}
no_vending = filter(subway_data, vending == "NO")
have_entry = filter(no_vending, entry == T)
```

+ The proportion of station entrances / exits without vending allow entrance is `r nrow(have_entry) / nrow(no_vending)`.

**Let's reformat the data!**

```{r}
tidy_station = distinct_station %>%
  mutate_at(vars(route1:route11), as.character) %>%
  pivot_longer(
    route1:route11,
               names_to = "route_name",
               values_to = "route_number")
  routenumberA = filter(tidy_station, route_number == "A")
  adacom = filter(routenumberA, ada == TRUE)
```

+ `r nrow(routenumberA)` distinct stations serve the A train

+ Of the stations that serve the A train, `r nrow(filter(adacom))` are ADA compliant.

## Problem 3

**Read and clean data in plos-monthin**

```{r}
poles_month_df = read.csv("~/R_columbia/p8105/p8105_hw2_xw2758/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate_at(vars(year:day), as.numeric) %>%
  mutate(month = recode(month, "1" = "January", "2" = "February", "3" = "March", "4" = "April", "5" ="May", "6" = "June", "7" = "July", "8" = "August", "9" = "September", "10" = "October", "11" = "November", "12" = "December")) %>%
  mutate(
    president = case_when(
      prez_gop == "1" ~ "gop",
      prez_dem == "1" ~ "dem"
  )) %>% 
  select(-prez_gop, -prez_dem, -day)
```

**Next, read and clean data in snp**

```{r}
snp_df = read.csv("~/R_columbia/p8105/p8105_hw2_xw2758/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate_at(vars(year:day), as.numeric) %>%
  mutate(month = recode(month, "1" = "January", "2" = "February", "3" = "March", "4" = "April", "5" ="May", "6" = "June", "7" = "July", "8" = "August", "9" = "September", "10" = "October", "11" = "November", "12" = "December")) %>% 
  select(-day) %>%
  relocate(year, month)
```

**Then, read and clean data in unemployment**

```{r}
unemployment_df = read.csv("~/R_columbia/p8105/p8105_hw2_xw2758/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unem_percentage") %>%
      mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" ="May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))
```


**Join three datasets**

```{r}
final_df = poles_month_df %>%
  left_join(snp_df, by = c("year","month")) %>%
  left_join(unemployment_df, by = c("month","year"))
```

**Last explain**

+ Pols-month has `r ncol(poles_month_df)` variables and `r nrow(poles_month_df)`observation. The range of year is `r range(pull(poles_month_df, year))`.

+ Snp has `r ncol(snp_df)` variables and `r nrow(snp_df)`observation. The range of year is `r range(pull(snp_df, year))`

+ Pols-month has `r ncol(poles_month_df)` variables and `r nrow(poles_month_df)`observation. The range of year is `r range(pull(poles_month_df, year))`.

+ Unemployment has `r ncol(unemployment_df)` variables and `r nrow(unemployment_df)`observation. The range of year is `r range(pull(unemployment_df, year))`.

+ Final result joint dataset has `r ncol(final_df)` variables and `r nrow(final_df)`observation. The range of year is `r range(pull(final_df, year))`. It has some key varibles, including `r names(final_df)`.